---

title: "Data Analytics in R - Assignment 1"
author: "Don Smith"
output: github_document

---

# Part 1: Questions 1 and 2: ACT scores

#### The ACT is a standardized college admissions test used in the United States. The four multi-part questions in this assessment all involve simulating some ACT test scores and answering probability questions about them.

#### For the three year period 2016-2018, ACT standardized test scores were approximately normally distributed with a mean of 20.9 and standard deviation of 5.7. (Real ACT scores are integers between 1 and 36, but we will ignore this detail and use continuous values instead.)

#### First we'll simulate an ACT test score dataset and answer some questions about it.

#### Set the seed to 16, then use rnorm() to generate a normal distribution of 10000 tests with a mean of 20.9 and standard deviation of 5.7. Save these values as act_scores. You'll be using this dataset throughout these four multi-part questions.

```{r}

set.seed(16)
act_scores <- rnorm(10000, 20.9, 5.7)

```

## Question 1a

#### What is the mean of act_scores?

```{r}

mean(act_scores)

```

## Question 1b

#### What is the standard deviation of act_scores?

```{r}

sd(act_scores)

```

## Question 1c

#### A perfect score is 36 or greater (the maximum reported score is 36).

#### In act_scores, how many perfect scores are there out of 10,000 simulated tests?

```{r}

length(act_scores[act_scores >= 36])

```

## Question 1d

#### In act_scores, what is the probability of an ACT score greater than 30?

```{r}

sum(act_scores > 30)/length(act_scores)

```

## Question 1e

#### In act_scores, what is the probability of an ACT score less than or equal to 10?

```{r}

mean(act_scores <= 10)

```

## Question 2
#### Set x equal to the sequence of integers 1 to 36. Use dnorm to determine the value of the probability density function over x given a mean of 20.9 and standard deviation of 5.7; save the result as f_x. Plot x against f_x.

```{r}

x = seq(1:36)

f_x <- dnorm(x, 20.9, 5.7)
plot (x, f_x)

```


# Part 2: Questions 3 and 4: ACT scores

#### In this 3-part question, you will convert raw ACT scores to Z-scores and answer some questions about them.

#### Convert act_scores to Z-scores. Recall from Data Visualization (the second course in this series) that to standardize values (convert values into Z-scores, that is, values distributed with a mean of 0 and standard deviation of 1), you must subtract the mean and then divide by the standard deviation. Use the mean and standard deviation of act_scores, not the original values used to generate random test scores.

```{r}

z_scores <- (act_scores-mean(act_scores))/sd(act_scores)
mean(z_scores > 2)

```

## Question 3b

#### What ACT score value corresponds to 2 standard deviations above the mean (Z = 2)?

```{r}
mean(act_scores) + (2*sd(act_scores))

```

## Question 3c

#### A Z-score of 2 corresponds roughly to the 97.5th percentile.

#### Use qnorm() to determine the 97.5th percentile of normally distributed data with the mean and standard deviation observed in act_scores.

#### What is the 97.5th percentile of act_scores?

```{r}

qnorm(.975, mean(act_scores), sd(act_scores))

```

#### In this 4-part question, you will write a function to create a CDF for ACT scores.



## Question 4a

#### What is the minimum integer score such that the probability of that score or lower is at least .95? Write a function that takes a value and produces the probability of an ACT score less than or equal to that value (the CDF). Apply this function to the range 1 to 36.

```{r}

cdf <- sapply(1:36, function (x){
  mean(act_scores <= x)
})
min(which(cdf >= .95))

```

## Question 4b

#### Use qnorm() to determine the expected 95th percentile, the value for which the probability of receiving that score or lower is 0.95, given a mean score of 20.9 and standard deviation of 5.7.

#### What is the expected 95th percentile of ACT scores?

```{r}

qnorm(0.95, 20.9, 5.7)

```

## Question 4c

#### We can use quantile() to determine sample quantiles from the data.

#### Make a vector containing the quantiles for p <- seq(0.01, 0.99, 0.01), the 1st through 99th percentiles of the act_scores data. Save these as sample_quantiles.

#### In what percentile is a score of 26?
#### (Your answer should be an integer (i.e. 60), not a percent or fraction. Note that a score between the 98th and 99th percentile should be considered the 98th percentile, for example, and that quantile numbers are used as names for the vector sample_quantiles.)

```{r}

sample_quantiles <- seq(0.01, 0.99, 0.01)
quantile(act_scores, sample_quantiles)

```

## Question 4d

#### Make a corresponding set of theoretical quantiles using qnorm() over the interval p <- seq(0.01, 0.99, 0.01) with mean 20.9 and standard deviation 5.7. Save these as theoretical_quantiles. Make a QQ-plot graphing sample_quantiles on the y-axis versus theoretical_quantiles on the x-axis.

```{r}

theoretical_quantiles <- qnorm(sample_quantiles, 20.9, 5.7)
qqplot(y=sample_quantiles, x=theoretical_quantiles)

```