---

title: "Data Analytics in R - Assignment 18"
author: "Don Smith"
output: github_document

---

# Part 1

### 1. Create a code chunk to complete the following tasks:

    Load the following packages: tidyverse, magrittr, lubridate, and corrplot. (You may need to install those packages if you have not already done so.)
    Read in the hmdaInterestRate.rds file.
    Report the structure of the dataframe. (No need to comment on the structure.)
    
```{r}

library(magrittr)
library(tidyverse)
library(lubridate)
library(corrplot)
library(tidyverse)
library(jtools)
library(ggstance)
library(huxtable)
library(corrplot)

```

### 2. Read in data. 

```{r}

df<- readRDS('C:\\Users\\Owner\\Documents\\GitHub\\r\\class assignments\\Data Analytics in R\\Assignment 18\\data\\hmdaInterestRate.rds')

str(df)

```


# Part 2

### Data preparation:

    Replace the values in the following columns with the same value divided by 1,000: loan_amount, property_value, and income. (This will make it easier to see the impact on the interest rate.)
    Create a new column, ltp, that is equal to the values in the loan_amount column divided by the values in the property_value column.
    Filter the data to keep observations for which income is less than 300 (i.e., $300,000).
    Report a summary of all columns. (No need to comment on the summary of the columns.)
    
```{r}

df1<-df%>% mutate(loan_amount= loan_amount/1000, property_value= property_value/1000, income= income/1000, ltp= loan_amount/property_value ) %>% filter(income<300)

summary(df1$income)

```


# Part 3

### Create a correlation plot of the following columns: interest_rate, ltp, income, applicant_age, property_value, and loan_amount.

    Below the plot, identify what variable has the strongest negative correlation with interest_rate. Comment on what might explain why that correlation is negative.

```{r}

ctrd<- cor(df1%>% select(where(is.numeric))) 
ctrd
corrplot(ctrd, method = 'color' , order = 'hclust', addCoef.col = 'black',number.cex = .6)

```

###### By looking at the plot, the variable that has the strongest negative correlation with interest rate is loan amount. This value (-0.33) represents a moderate negative correlation.  This makes sense because higher loan amounts are typically taken by wealthier individuals who are low-risk for the lender, so the lender gives a lower interest rate.  It's worth noting that interest rate and "ltp" also had a moderate negative correlation (-0.32)


# Part 4

### Regress interest rate on ltp. Interpret the coefficient estimate on ltp.

```{r}

Model1<- lm(interest_rate~ltp, data=df1) 
summary(Model1)

```


```{r}

ggplot(df1, aes(x = ltp, y = interest_rate)) +
  geom_point() +
  expand_limits(x = c(0,1)) +
  stat_smooth(method = 'lm', fullrange = T)

```

###### As ltp increases by one unit, the interest rate decreases by 1.05.


# Part 5

### Regress interest rate on ltp and loan_amount. Comment on the change in the adjusted R-squared, as well as the change in the coefficient on ltp.

```{r}

Model2<- lm(interest_rate ~ltp +loan_amount, data=df1)
summary(Model2)
export_summs(Model1, Model2)

```


```{r}

plot_summs(Model1, Model2)

```

###### The change in coefficient is -1.05 for model 1 and -.74 for model 2 this means that model 1 has a stronger negative correlation than model 2.The difference between the two models in adjusted r squared is .1 for model 1 and .1522 for model 2. Model 2 can now explain more variation in interest rate than it could before. Model 1 can only explain 10% of the variation in interest rate.


# Part 6

### Regress interest rate on ltp, loan_amount, and aus_1. Interpret the new coefficients.
```{r}

Model3<- lm(interest_rate~ltp +loan_amount +aus_1, data=df1)
summary(Model3)
export_summs(Model1, Model2, Model3)

```


```{r}

plot_summs(Model1, Model2, Model3)

```

###### As we add model 3 to the regression, the r^2 value increased from .1 to .15 to .27. This tells us that model 3 now accounts for 27% of the variation in interest rate. As we add more significant predictive variables, we expect the r^2 value to increase.
