---

title: "Data Analytics in R - Assignment 8"
author: "Don Smith"
output: github_document

---

### Question 1: Read in the stars dataset, which is included in base R. Include any packages you need to do your analysis.

```{r}

library(tidyverse)
library(dslabs)
library(ggplot2)

data(stars)

```

### Question 2: Review the first 6 rows of the dataset.

```{r}

head(stars)

```

### Question 3: What are the mean and standard deviation of the magnitude column?

```{r}

mean(stars$magnitude)
sd(stars$magnitude)

```

### Question 4: Create a density plot using the magnitude column.

```{r}

ggplot(stars, aes(x=magnitude)) + geom_density(color="darkblue", fill="lightblue")

```

### Question 5: Plot the relationship between magnitude and tempurature.

```{r}

stars %>%  ggplot(aes(temp, magnitude, color = type)) + geom_point()

```

### Question 6: Plot the relationship between magnitude and tempurature, this time with y in the logarithmic scale.

```{r}
stars %>%  ggplot(aes(temp, magnitude)) + geom_point() + scale_y_reverse() + scale_y_continuous(trans = "log10") + scale_x_reverse()

```

### Question 7: How many stars have a magnitude greater than 10 and temperature greater than 5000?

```{r}

s1 <- stars %>% filter(magnitude >10 & temp >5000)
s1

count(s1)

```

### Question 8: How many stars have a stars with magnitude less than 0 and temperature less than 3500?

```{r}

s2 <- stars %>% filter(magnitude < 0 & temp < 3500)
s2

count(s2)

```


