---

title: "Data Analytics in R - Assignment 9"
author: "Don Smith"
output: github_document

---
## Question 1: Read in Titanic data.

```{r}

library(dplyr)
library(ggplot2)
library(dslabs)
library(tidyverse)

#Read in data
#Note: For the purposes of this assignment, data has been truncated
titanic_train <- read.csv("C:\\Users\\Owner\\Documents\\GitHub\\r\\class assignments\\Data Analytics in R\\Assignment 09\\data\\titanic_data.csv")
head(titanic_train)

```

## Question 2: Filter data set to isolate columns c(Survived, Pclass, Sex, Age, SibSp, Parch, Fare).

```{r}
#Isolated requisite columns, and converted Survived, Pclass and Sex to factors for future use
titanic <- titanic_train %>%
  select(Survived, Pclass, Sex, Age, SibSp, Parch, Fare) %>%
  mutate(Survived = factor(Survived),
         Pclass = factor(Pclass),
         Sex = factor(Sex))

head(titanic)

```
## Question 3: Determine the number of survivers vs non-survivers aggregated by class.

```{r}

#Counting number of survived/died by class
survived1 <- count(titanic %>% filter(Pclass == 1, Survived == 1))
died1 <- count(titanic %>% filter(Pclass == 1,Survived == 0))

survived2 <- count(titanic %>% filter(Pclass == 2, Survived == 1))
died2 <- count(titanic %>% filter(Pclass == 2,Survived == 0))

survived3 <- count(titanic %>% filter(Pclass == 3, Survived == 1))
died3 <- count(titanic %>% filter(Pclass == 3,Survived == 0))


# Loading data into data.frame for easier viewing
table_data <- data.frame(
  Row_names = character(),
  FirstClass = integer(),
  SecondClass= integer(),
  Thirdclass= integer())

table_data[1,] <- c('Survived', as.integer(survived1), as.integer(survived2), as.integer(survived3))
table_data[2,] <- c('Died', as.integer(died1), as.integer(died2), as.integer(died3))

table_data


```


## Question 4: Create a density plot showing survivers vs non-survivers.

```{r}

titanic %>%  ggplot(aes(Age, fill = Survived)) +
  geom_density(aes(y = ..count..), alpha = 0.2, bw = 0.5, position = "stack")

```


## Question 5: Create a density plot showing survivers vs non-survivers, aggregated by class, gender and age.

```{r}

titanic %>%  ggplot(aes(Age, fill = Survived)) +
  geom_density(aes(y = ..count..), alpha = 0.2, bw = 0.5, position = "stack") + facet_grid(Sex~Pclass)

```

## Question 6: Create a bar graph showing the number of survivers vs non-survivers, aggregated by class.

```{r}

titanic %>% ggplot(aes(Pclass, fill = Survived)) + geom_bar()

```

## Question 7: Create a box plot of survivers vs non-survivers, aggregated by fare price.

```{r}

titanic %>% filter(Fare >0) %>% 
  ggplot(aes(Survived, Fare, fill = Survived)) + scale_y_continuous(trans = "log2") + geom_boxplot() + geom_jitter(width = 0.1, alpha = 0.2) 

```

